|

### 2021-12-15 02-Chp-Systematic-Review-Tools.Rmd
Update table captions to new style
Ciara edits

# medrxivr: an R package for systematically searching biomedical preprints {#sys-rev-tools-heading}

@@

Preprints are copies of academic manuscripts that are posted online in advance of being formally published by an academic journal. They represent an important source of scientific literature. A new software program called `medrxivr` was created as part of this thesis to allow researchers to find preprints related to their research in a transparent and reproducible way. Development of this tool was an essential part of this thesis, as preprints represent a key source of information needed for the research reported in future chapters.

@@

Preprints represent an increasingly important source of scientific information (see Section \@ref(diverse-sources-preprints)). As a result, repositories of preprinted articles should be considered a distinct but complementary information source when reviewing the evidence base as part of a systematic review. The two key repositories in the health science are bioRxiv, established in 2013,[@sever2019] and medRxiv, which launched in 2019 and was designed to replace the "Epidemiology" and "Clinical Trial" categories of bioRxiv.[@rawlinson2019]

Searching these preprints as part of the systematic review described in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

This chapter outlines the development and key functionality of `medrxivr` (version 0.0.5), a tool I created to facilitate the systematic searching of medRxiv and bioRxiv preprints. The factors that necessitated the development of this tool in the context of this thesis are outlined, and the use of `medrxivr` in my own projects and by other researchers is discussed. As the majority of work on this aspect of my thesis is represented by lines of code or online documentation (available at [https://github.com/ropensci/medrxivr](https://github.com/ropensci/medrxivr) and [https://docs.ropensci.org/medrxivr/](https://docs.ropensci.org/medrxivr/) respectively), this chapter is an intentionally short, high-level summary of my work on this project. The GitHub repository for the `medrxivr` contains a complete record of the development of this tool, including discussion with other members of the systematic review community.[@zotero-15029]

@@

```{r medrxivr-sr, echo = FALSE, fig.align='center',results="asis", fig.cap='(ref:medrxivr-sr-cap)', out.width='65%', fig.scap='(ref:medrxivr-sr-scap)'}
knitr::include_graphics(file.path("figures","sys-rev-tools","medrxiv-role.png"))
```

@@

I developed the tool to meet three success criteria,[@wateridge1995] influenced both by the functionality required to perform systematic searches as part of the review presented in the subsequent chapters, discussion with information specialist colleagues, and an informal survey of the evidence synthesis and health librarian communities on Twitter. The criteria were as follows:

@@

2.  support for bulk export of references returned by the search to a file type that can be readily imported into a reference manager (e.g., *.bib* or *.ris*); and

3.  automated retrieval of the full-text PDFs of relevant records, similar to the Find Full Text feature offered by EndNote.

@@

Prior to development of this tool, I conducted an audit of existing tools for accessing medRxiv and bioRxiv metadata. While none addresses the success criteria described above, two of these tools are useful to consider to highlight the additional functionality that `medrxivr` contributes.

The first, a platform called Rxivist[@abdill2019], allows users to search preprints using keywords. However, the core functionality of the Rxivist platform is focused around exploring the number of times a preprint has been downloaded and/or shared on Twitter, to allow researchers to find the most popular papers related to their topic. The search interface[@zotero-15027] does not allow for complex search strategies using Boolean operators and there is no option to batch-export the results of a search.

The second tool, `search.bioPreprint`, allows users to search for terms across a range of preprint servers, including medRxiv and bioRxiv, but also journals which use a post-publication peer-review process such as F1000Research.[@iwema2016] However, similar to the Rxivist platform, this tool is designed for researchers aiming to keep up to date with recent developments in their fields rather than systematically assess the entirety of the available literature. As such, the platform only returns the most recent `r comma(1000)` records by publication date.

@@

Work on the `medrxivr` tool began in Summer 2019, and initially consisted of a development of set of R scripts to allow for searching medRxiv and bioRxiv as part of the systematic search outlined in Chapter \@ref(sys-rev-methods-heading). Following interest from other researchers in using the *ad-hoc* web-scraping scripts, additional development work took place in 2019/2020, allowing for improved searching and exporting functionality and I released the initial version of the `medrxivr` R package in February 2020.

Early versions of the tool had a reliance on scraping data directly from the repository website. Web-scraping is a fragile mechanism for extracting data, as it is entirely dependent on consistent website design and underlying code structure remaining unchanged.[@shaw2002; @laprie1992]. In the case of `medrxivr`, as the medRxiv/bioRxiv websites are regularly updated, ensuring the web-scraping performed as expected required that I regularly update or fix the script.

@@

While the package allows users to interact with and search both medRxiv and bioRxiv metadata, as the process is identical for both, searching the medRxiv database is used as an illustrative example throughout this chapter.

@@

Prior to searching the metadata, it must first be imported in R. In `medrixvr`, I have provided two separate but related methods for users to import the data (Figure \@ref(fig:medrxivr-data-sources)). The first of these methods, accessed via the `mx_api_content()` function, creates a local copy of all data available from the medRxiv API at the time the function is run.

@@

This provides an up-to-the-minute reflection of the medRxiv preprint repository. However, this approach has two limitations. Firstly, as the API returns results as a series of pages limited to 100 records per page, downloading the entire database requires a time-intensive process of cycling through each page. Secondly, the API can become unavailable, either during peak usage times or planned maintainence windows. To address these limitations, I provide a second method of accessing medRxiv data, called via the `mx_snapshot()` function, which allows users to access a maintained static snapshot of the database.

@@

This snapshot is created each morning at 6am using a process known as "git-scraping",[@zotero-15031] whereby the entire database is downloaded using the `mx_api_content()` function and saved as a comma separated value (CSV) file in an online repository (Figure \@ref(fig:medrxivr-data-sources)). Calling `mx_snapshot()` imports this CSV into R, and has the advantage of both faster loading of the data into R (as it is imported as a single file and does not require cycling through the output of the API) and an absence of any reliance on the API. The one limitation of this approach is that the snapshot (by its nature) will not contain details of records added to the database since it was taken. However, given that the number of records added each day is relatively low, this should pose minor issues.

@@

@@

An important argument of the `mx_search()` is `report`, which outputs a structured table with each search strategy presented on an individual line and the number of records associated with this strategy.[@rethlefsen2021prisma]

@@

This allows users to discover which terms in their search are contributing most to the total number of results returned. This is important as part of developing a search strategy,[@bramer2018] as it allows for the key terms related to each topic to be discovered. It also aids in identifying misspelled or case-sensitive search terms, which will frequently return no results. As an example, in the search presented above, the term "alzheimer's" returns no records. This is expected, as "Alzheimer's" is a proper noun and so should be capitalised, but serves to illustrate the usefulness of the reporting function.

In hindsight, the decision to make the search case-sensitive was a poor one as users reported being unaware of this setting when searchin. However, changing the default behaviour of `mx_search()` to be case-insensitive is problematic, as it could result in different numbers of hits based solely on which version of the package the user has installed. To address this, `mx_search()` now contains an `autocaps` argument, which when set to `TRUE` will automatically search for both capitalised and uncapitalised versions of the search term.

@@

In line with my second success criteria (Section \@ref(success-criteria)), one of the key features of the `medrxivr` is the ability for users to export easily the results of their systematic search to a reference manager. While it is a seemingly simple request, this is is one of the key ways in which `medrxivr` is set apart for other preprint search tools, including the native medRxiv/bioRxiv website search functionality.

For example, the results of our simple search above can be exported to the `"medrxiv_export.bib"` file using the following code:

@@

`medrxivr` also allows users to download the full text papers for records that are deemed eligible for full-text screening (see Figure \@ref(fig:medrxivr-sr)). `mx_download()` takes the list of included records and saves the PDF for each to a folder specified by the user. This functionality is similar to the "Find Full Text" feature offered by EndNote.

@@

Lobbying of the Cold Springs Harbor Laboratory to develop the API to allow for direct searching of the database has been ongoing. This would negate the current need to download a local copy of the relevant preprint database before searching it, which is currently the rate limiting step for performing searches. For example, as of January 2021, downloading a copy of the bioRxiv database takes approximately an hour.

@@

In addition to being used to search systematically health-related preprint servers, as illustrated in the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading), `medrxivr` has other uses.

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-tools-pub)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a subtantial number of preprints are never formally published but remain accessible as preprints.

@@

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis, there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, and so some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

There is also the potential that the cross-section of literature posted on medrxiv/bioRxiv is substantially different to those suffering from publication bias (studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant).[@song2010] This is because simply lowering the barriers to publication may well encourage authors to published "null" results, but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[@rosenthal1979]

@@

-   I have contrasted `medrxivr` with other available interfaces to medRxiv/bioRxiv data to highlight the added functionality it offers. I have also discussed the tools reception to date, its limitations, and the important role of open-source tools like `medrxivr` in evidence synthesis.

>>>>

# `medrxivr`: an R package for systematically searching biomedical preprints {#sys-rev-tools-heading}

@@

Preprints are copies of academic manuscripts that are posted online in advance of being formally peer-reviewed and published by an academic journal. As part of this thesis, a new software program called `medrxivr` was created to allow researchers to find preprints related to their research in a transparent and reproducible way. Development of this tool was important, as preprints represent a key source of information needed for the research reported in future chapters.

@@

Preprints represent an increasingly important source of scientific information (see Section \@ref(diverse-sources-preprints)). As a result, when reviewing the evidence base as part of a systematic review, repositories of preprinted articles should be considered a distinct but complementary information source to databases of published articles. The two key repositories in the health science are bioRxiv, established in 2013,[@sever2019] and medRxiv, which launched in 2019 and was designed to replace the "Epidemiology" and "Clinical Trial" categories of bioRxiv.[@rawlinson2019]

Searching these preprints as part of the systematic review described in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading) was a necessity because many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID-19, this already time-consuming and error-prone method is no longer feasible.

This chapter outlines the development and key functionality of `medrxivr` (version 0.0.5), a tool I created to facilitate the systematic searching of medRxiv and bioRxiv preprints. The factors that necessitated the development of this tool in the context of this thesis are outlined, and the use of `medrxivr` in my own projects and by other researchers is discussed. As the majority of work on this aspect of my thesis is represented by lines of code or online documentation (available at [https://github.com/ropensci/medrxivr](https://github.com/ropensci/medrxivr) and [https://docs.ropensci.org/medrxivr/](https://docs.ropensci.org/medrxivr/) respectively), this chapter is an intentionally short, high-level summary of my work on this project. The GitHub repository for `medrxivr` contains a complete record of the development of this tool, including discussion with other members of the systematic review community.[@zotero-15029]

@@

```{r medrxivr-sr, echo = FALSE, fig.pos="h", fig.align='center',results="asis", fig.cap='(ref:medrxivr-sr-cap)', out.width='65%', fig.scap='(ref:medrxivr-sr-scap)'}
knitr::include_graphics(file.path("figures","sys-rev-tools","medrxiv-role.png"))
```

@@

I developed the tool to meet three success criteria:[@wateridge1995]

@@

2.  support for bulk export of references returned to a file type that can be readily imported into a reference manager (e.g., *.bib* or *.ris*); and

3.  automated retrieval of the full-text PDFs of relevant records.

These criteria were influenced by the need for functionality to perform systematic searches as part of the review presented in the subsequent chapters, discussion with information specialist colleagues about key features, and an informal survey of the evidence synthesis and health librarian communities on Twitter.

@@

Prior to development of this tool, I conducted an audit of existing tools for accessing medRxiv and bioRxiv metadata. While none addresses the success criteria described above, two of these tools are useful in highlighting the additional functionality that `medrxivr` contributes.

The first, a platform called Rxivist[@abdill2019], allows users to search preprints using keywords. However, the core functionality of the Rxivist platform is focused around exploring the number of times a preprint has been downloaded and/or shared on Twitter, to allow researchers to find the most popular papers related to their topic. Additionally, the search interface[@zotero-15027] does not allow for complex search strategies using Boolean operators.

The second preexisting tool, `search.bioPreprint`, allows users to search for terms across a range of preprint servers, including medRxiv and bioRxiv, but also journals which use a post-publication peer-review process such as F1000Research.[@iwema2016] However, similar to the Rxivist platform, this tool is designed for researchers aiming to keep up to date with recent developments in their fields rather than systematically assess the entirety of the available literature. As such, the platform only returns the most recent `r comma(1000)` records by publication date.

@@

Work on the `medrxivr` tool began in Summer 2019, and initially consisted of a development of a set of R scripts to allow for searching medRxiv and bioRxiv as part of the systematic search outlined in Chapter \@ref(sys-rev-methods-heading). Following interest from other researchers in using the *ad-hoc* web-scraping scripts, additional development work took place in 2019/2020, allowing for improved searching and exporting functionality. I released the initial version of the `medrxivr` R package in February 2020.

Early versions of the tool had a reliance on scraping data directly from the repository website. Web-scraping is a fragile mechanism for extracting data, as it is entirely dependent on consistent website design and underlying code structure remaining unchanged.[@shaw2002; @laprie1992]. In the case of `medrxivr`, as the medRxiv/bioRxiv websites are regularly updated, ensuring the web-scraping script continued to perform as expected required substantial maintenance work.

@@

While the package allows users to interact with and search both medRxiv and bioRxiv metadata, because the process is identical for both, searching the medRxiv database is used as an illustrative example throughout the remainder of this chapter.

@@

Prior to searching the metadata, users must first create a local copy of the database, denoted using `mx_data` below. In `medrixvr`, I have provided two separate but related methods to achieve this (Figure \@ref(fig:medrxivr-data-sources)). The first of these methods, accessed via the `mx_api_content()` function, creates a local copy of all data available from the medRxiv API at the time the function is run.

@@

This provides an up-to-the-minute reflection of the medRxiv preprint repository. However, this approach has two limitations. Firstly, because the API returns results as a series of pages limited to 100 records per page, downloading the entire database requires a time-intensive process of cycling through each page. Secondly, the API can become unavailable, either during peak usage times or planned maintainence windows. To address these limitations, I provide a second method of accessing medRxiv data, called via the `mx_snapshot()` function, which allows users to access a maintained static snapshot of the database. Note that due to the size of bioRxiv, only a maintained snapshot of the medRxiv repository is available via `mx_snapshot()`.

@@

This snapshot is created each morning at 6am using a process known as "git-scraping",[@zotero-15031] whereby the entire database is downloaded using the `mx_api_content()` function and saved as a comma separated value (CSV) file in an online repository (Figure \@ref(fig:medrxivr-data-sources)). Calling `mx_snapshot()` imports this CSV into R, and has the advantage of both faster loading of the data into R (because it is imported as a single file and does not require cycling through the output of the API) and an absence of any reliance on the API. The one limitation of this approach is that the snapshot (by its nature) will not contain details of records added to the database since it was taken. However, given that the number of records added each day is relatively low, this should only pose minor issues.

@@

Note that for my example, I took a two-step approach by first creating the full search term `myquery` which is then passed to `mx_search()`. This is not necessary to perform a search, as users can pass terms directly to `mx_search()`:

```{r, eval = FALSE}
# Run a search to find all records with "dementia"
results <- mx_search(mx_data,
                     "dementia")
```

@@

An important feature of the `mx_search()` is the `report` functionality, which outputs a structured table showing each search term along with the corresponding number of records.[@rethlefsen2021prisma]

@@

This is important as part of developing a search strategy,[@bramer2018] because it allows for the key terms related to each topic to be discovered. It also aids in identifying misspelled or case-sensitive search terms, which will frequently return no results. As an example, in the search presented above, the term "alzheimer's" returns no records. This is expected, because "Alzheimer's" is a proper noun and so should be capitalised, but serves to illustrate the usefulness of the reporting function.

In hindsight, the decision to make the search case-sensitive was a poor one because users reported being unaware of this setting when carrying out their search. However, changing the default behaviour of `mx_search()` to be case-insensitive is problematic because it could result in different numbers of hits based solely on which version of the package the user has installed. To address this, `mx_search()` now contains an `autocaps` argument, which when set to `TRUE` will automatically search for both capitalised and uncapitalised versions of the search term.

@@

In line with my second success criteria (Section \@ref(success-criteria)), one of the key features of the `medrxivr` is the ability for users to export easily the results of their systematic search to a reference manager. While this would seem a likely candidate to be a generic feature, this is one of the key ways in which `medrxivr` is set apart from other preprint search tools. None of tools that I audited provide this feature, nor does the native medRxiv/bioRxiv website search.

The results of our simple search above can be exported to the `medrxiv_export.bib"` file using the following code:

@@

`medrxivr` also allows users to download the full text papers for records that are deemed eligible for full-text screening (see Figure \@ref(fig:medrxivr-sr)). `mx_download()` takes the list of included records and saves the PDF for each to a folder specified by the user.

@@

Lobbying of the Cold Springs Harbor Laboratory to develop the API to allow for direct searching of the database has been ongoing. This would negate the current need to download a local copy of the relevant preprint database before searching it, which is currently the rate limiting step for performing searches. For example, as of January 2021, downloading a copy of the bioRxiv database takes approximately one hour.

@@

In addition to being used to systematically search health-related preprint servers, as illustrated in the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading), `medrxivr` has other uses.

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-tools-pub)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a substantial number of studies are never formally published but remain accessible as preprints.

@@

While searching the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis, there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, and so some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

There is also the potential that the cross-section of literature posted on medRxiv/bioRxiv is substantially different from those suffering from publication bias (studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant).[@song2010] This is because simply lowering the barriers to publication may well encourage authors to published "null" results, but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[@rosenthal1979]

@@

-   I have contrasted `medrxivr` with other available interfaces to medRxiv/bioRxiv data to highlight the added functionality it offers. I have also discussed the tool's reception to date, its limitations, and the important role of open-source tools like `medrxivr` in evidence synthesis.
